Learning constraints in spreadsheets and tabular data

### 2016-12-05 Paper/response_letter/response_letter.tex

\documentclass{article}
\usepackage{enumerate}
\usepackage[english]{babel}
\usepackage{times}
\usepackage{hyperref}
\usepackage{latexsym}    % to get LASY symbols
\usepackage{graphicx}    % to insert PostScript figures
\usepackage{xspace}
\usepackage[usenames,dvipsnames]{xcolor}

\newcommand{\sergey}[1]{\textcolor{magenta}{{\sc Sergey:} #1}\xspace}
\newcommand{\sam}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
\newcommand{\samuel}[1]{\textcolor{green}{{\sc Samuel:} #1}\xspace}
\newcommand{\tias}[1]{\textcolor{blue}{{\sc Tias:} #1}\xspace}
\newcommand{\luc}[1]{\textcolor{red}{{\sc Luc:} #1}\xspace}
\newcommand{\added}[1]{\textcolor{red}{added:} {#1} \textcolor{red}{end}\xspace}

\newcommand{\scaleconstant}{1.0}
\bibliographystyle{plainnat}
\author{Samuel Kolb \and Sergey Paramonov \and Tias Guns \and Luc {De Raedt}}
\title{MLJ Revision Letter:\\ learning constraints in spreadsheets and tabular data}
\begin{document}
\maketitle
\sergey{We need a discussion section on: noise, (non)-contiguous blocks and combination of constraints} 
\sergey{just in case, the text is preliminary, feel free to cut, expand and rewrite, I think it should be 1 or 2 pages at the end, I hope}
\section{Key issues}

\paragraph{Goal and motivation}
\sergey{Reviewer pointed out that (and we should thank him and use it, I think):}

\textbf{Reviewer \#3:}
Section 6. The reviewer would like to point out the important field of data cleaning (of dirty data, as it is called in the database community) and anomaly detection, beyond simple error detection. Because your method does not only detect errors, but can also show *why* it deems something as erronous your method falls in the desirable category of descriptive or explainable anomaly detection algorithms.\\
\sergey{A sentence on compression, maybe?}\\
\sergey{We need to connect better Sections 5 and 6 -- from the reviews}

@@

\paragraph{Approach}
\sergey{here is the response on the unclear definitions and notation}

\paragraph{Competitor algorithms and related work:} \textit{The reviewers point out that you should compare with other methods. One reviewer stresses that Microsoft Excel provides functionality that alleviates part of the problem you address. Another reviewer suggests that you compare to a baseline that considers all constraints. You must compare to algorithms that solve the same problem (at least part of it).}

\sergey{Tias, you also had a point here that flash fill is searching for a combination of string transformations, we search for atomic constraints and etc, could you integrate it here? and indicate what part should be added to the paper?}

### 2016-12-09 Paper/ECAI/tabular.tex
Reworked motivation

\section{Introduction}
% relevance of formulas and spreadsheets
Millions of people across the world use spreadsheets every day.
The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
Many end-users lack the understanding of the underlying structures and relations in such sheets and the data they contain.
This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
In this case, often a comma-separated values (CSV) format is used meaning that all formulas are lost, including inter-sheet formulas and relations.
%This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across all tables.
For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulae.

% why learn them
Such issues and mistakes could be overcome through the development of intelligent user-assisting tools. A core aspect of such tools would be the ability to automatically discover what constraints and formulae are present in a spreadsheet. This is the problem we study in this paper. There are multiple applications that could build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.

@@

% Moved up
The question that we ask in this paper is: is it possible to learn the set of constraints present in a spreadsheet, from just flat tabular spreadsheet data? The main challenge is the number of possible constraints and combinations of rows and columns that need to be tried as input to the constraints.
To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
%
% contribs
Our contributions are as follows:
\begin{itemize}
\item we define the tabular constraint learning problem, where the goal is to find constraints that range over entire rows or columns in an unsupervised way;
\item we propose an effective 2-stage generate and test method where the first stage reasons only over properties of contiguous blocks of rows/columns, and the second stage continues to investigate individual rows and columns and their content;
\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximally contiguous blocks compatible with the candidate constraint's argument; %, using backtracking search and forward checking;
\item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall. 
\end{itemize}

@@

This paper is organized as follows.
Section \ref{sec:related_work} discusses related work. Section \ref{sec:formalization} introduces the relevant concepts and defines the problem formally while Section \ref{sec:approach} presents our approach. This is evaluated in Section \ref{sec:evaluation} after which we show two applications of how our system could be used in Section \ref{sec:applications}, and we conclude in Section \ref{sec:conclusions}.

>>>>

\section{Introduction}
% relevance of formulas and spreadsheets
Millions of people across the world use spreadsheets every day.
The tabular representation of the data is often intuitive, and the programming of functions in individual cells is quickly learned.
However, large and complex sheets (possibly with multiple tables and relations between them) can be hard to handle.
Many end-users lack the understanding of the underlying structures and relations in such sheets and the data they contain.
This is especially the case when spreadsheets have been exported from other software such as Enterprise Resource Planning (ERP) systems.
In this case, often a comma-separated values (CSV) format is used meaning that all formulas are lost, including inter-sheet formulas and relations.
%This limited understanding is especially the case with spreadsheets exported from other software such as ERP packages. In this case,
Even in manually created spreadsheets, it can be challenging to be consistent and correct with formulas across all tables.
For example the influential Reinhart-Rogoff economical paper ``Growth in a Time of Debt'' \cite{growth_in_time_of_debt} had some of its claims contested~\cite{flaw_excel}, after an investigation of the used Excel sheets was shown to contain mistakes in formulas.

% why learn them
\added{
Such issues and mistakes could be overcome through the development of intelligent user-assisting tools.
A core aspect of such tools would be the ability to automatically discover what constraints and formulas are present in a spreadsheet.
This is the problem we study in this paper.
Concretely we consider five key applications that intelligent tools based on automatic constraint discovery may facilitate: auto-completion, error checking, formula suggestion, rich import and data compression.
Auto-completion helps users quickly populate spreadsheets by suggesting values for unfilled cells based on constraints that have previously been detected within the spreadsheet.
Error checking helps uncover erroneous values in spreadsheets, for example, while typing, by finding previously discovered constraint that are violated by new values.
Both of these applications can facilitate the use of spreadsheets even for users who do not explicitly use formulas at all.
Of course, these applications are also able to explain to more advanced users what constraints are responsible for suggesting or rejecting a value.
Formula suggestion is a variant of auto-completion that, instead of suggesting a \emph{value}, provides a formula that computes a user provided value.
This setting can help users to find formulas that perform a certain task or find the right syntax for a formula.
We use rich import to denote the task of importing semi-structured data from, for example, CSV files and reconstruct constraints that were used, implicitly or explicitly, to generate the data.
Rich import goes hand-in-hand with data compression, since a file can be compressed by identifying all values that are generated by a formula and storing that formula instead of the values generated by it.
%There are multiple applications that could build on this, such as rediscovery of constraints when importing raw data, auto-suggestion of formula's during spreadsheet use, automatic auto-completion of rows or columns, and consistency and error checking.
}

@@

% Moved up
The question that we ask in this paper is: is it possible to learn the set of constraints present in a spreadsheet, from just flat tabular spreadsheet data? The main challenge is the number of possible constraints and combinations of rows and columns that need to be tried as input to the constraints.
To answer this challenge we propose a general-purpose method and system, named \sname (from: Tabular Constraint Learner, pronounced ``tackle''), for discovering row-wise and column-wise constraints.
%
% contribs
Our contributions are as follows:
\begin{itemize}
\item we define the tabular constraint learning problem, where the goal is to find constraints that range over entire rows or columns in an unsupervised way;
\item we propose an effective 2-stage generate and test method where the first stage reasons only over properties of contiguous blocks of rows/columns, and the second stage continues to investigate individual rows and columns and their content;
\item furthermore, in the first stage we use a constraint solver to efficiently enumerate all combinations of maximally contiguous blocks compatible with the candidate constraint's argument; %, using backtracking search and forward checking;
\item experiments on different publicly available spreadsheets show that the system is able to extract constraints with high precision and recall.
\added{\item we describe on a high level how auto-completion and error checking can be realized using \sname}
\end{itemize}

@@

This paper is organized as follows.
Section \ref{sec:related_work} discusses related work. Section \ref{sec:formalization} introduces the relevant concepts and defines the problem formally while Section \ref{sec:approach} presents our approach. This is evaluated in Section \ref{sec:evaluation} after which \added{we show how to realize two of the motivating applications using our system} in Section \ref{sec:applications}, and we conclude in Section \ref{sec:conclusions}.
